---
title: "class_09"
author: "Jiaxin Qian"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Analysis of Human Breast Cancer Cells 
1. Input and organize the data 
```{r}
url <- "https://bioboot.github.io/bimm143_W18/class-material/WisconsinCancer.csv"
wisc.df <- read.csv(url) #ignore the funkey thingd in the data set, including ID, ddiagnosis, and the last X
wisc.data <- as.matrix(wisc.df[,3:32])
# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id
#head(wisc.data)
diagnosis <- wisc.df[["diagnosis"]]

```

> Q1. How many observations are in this dataset?

```{r}
nrow(wisc.df)
```

> Q2. How many of the observations have a malignant diagnosis?

```{r}
table(wisc.df$diagnosis)
```

> Q3. How many variables/features in the data are suffixed with _mean?

```{r}
#grep("_mean", colnames(wisc.df), value = TRUE)
length(grep("_mean", colnames(wisc.df)))
```

## PCA
Before we do PCA, we need to think or consider whether we should SCALE our input
It is important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data include:
**The input variables use different units of measurement.**
**The input variables have significantly different variances.**

```{r}
round(apply(wisc.data, 2, sd), 2) 
# check for large values
```

```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data, scale. = TRUE)
summary(wisc.pr)
```

##Interpreting PCA
```{r}
bipolt(wisc.pr)
#too messy to be understandable
```

```{r}
attributes(wisc.pr)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis,  xlab = "PC1", ylab = "PC2")
abline(h=0, col="grey", lty=2)
abline(v=0, col="grey", lty=2)
#red for M
```

###Cluster in PC space
cluster the original data
```{r}
hc <- hclust(dist(wisc.data))
plot(hc)
```

cluster the PCA
```{r}
wisc.pr.hc <- hclust(dist(wisc.pr$x[, 1:3]), method="ward.D2")
plot(wisc.pr.hc)
```

We can use `cutree()` to seperate the data 
```{r}
table(cutree(wisc.pr.hc, k=4))
gp3 <- cutree(wisc.pr.hc, k=4)
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=gp3)
```

We can use `table()` to compare the $diagnosis vector with out cluster result
```{r}
table(gp3, wisc.df$diagnosis)
```

##Prediction
```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
#use the predict() function with our previous PCA model and new data
npc <- predict(wisc.pr, newdata=new)
npc
```

now draw the PCA plot again and add our new data
```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=wisc.df$diagnosis)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```

